Below-ground sink limitation alters growth and carbon balance of Eucalyptus seedlings

### 2015-12-01 manuscript/manuscript_pve.Rmd
abbreviations updated

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, C allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (A) to allocate daily C uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in A explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in down regulation of A, these reductions were not large enough to explain observed growth responses. Thus, as A and growth were not tightly coordinated, excess photosynthetic C not attributed to biomass resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirms that important feedbacks exist between belowground sink strength and leaf C uptake.

@@

Understanding plant growth and its relationship to C assimilation requires knowledge of the mass balance that must be achieved between C uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth should principally depend on the allocation of photosynthate among different tissues and organs. At long enough time scales photosynthesis (A) and respiratory losses together determine net C balance and will be coordinated to plant growth. At shorter temporal scales, however, growth can instead be mediated by tissue C storage pools. This has led to the current debate on how strongly plant growth is controlled by either source or sink activity. Consequently, plant growth cannot always be simply determined by the photosynthesis rate, making it complex to understand and challenging to model [@fourcaud2008plant]. Despite a wealth of studies, large uncertainties still remain regarding the coordination of C supply and growth of woody species.

\
In woody species, the coordination of A and growth has been studied with manipulations of C source activity. Examples included elevated CO~2~ experiments, for example FACE [reviewed in @ainsworth2005have], and partial defoliation experiments. Elevated CO~2~ has been shown to increase A [@drake1997more; @ainsworth2007response] and across four FACE experiments this resulted in a stimulation of 23 % in forest biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. In partial defoliation experiments, increases in A of the remaining foliage are commonly shown, yet are attributed to various mechanisms, including reduction in end product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end], increased stomatal conductance [@layne1995end], enhanced leaf nutrient status [@turnbull2007increased] and regulatory sugar signaling [@eyles2013whole]. However, increases in A in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], C limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant C gain [@ovaska1993a]. These manipulations of C source activity expose unresolved issues with how changes in A do not always infer similar responses in growth.

\
Alternatively, manipulating plant tissue C sinks is often used to investigate the correlation of A and growth. This is because metabolic signaling networks, relaying information on C and N status of different tissues, can regulate photosynthetic activity [@paul2001sink]. If sink inhibition of A occurs, a close coordination between declines in A and growth should be expected. Whether photosynthetic down regulation is evident in woody species has been tested through fruit removal, phloem girdling and low temperatures at high elevations. In these studies, down regulation of A was frequently correlated to carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@hoch2002altitudinal; @urban2007girdling; @haouari2013fruit]. However, reductions in A were also attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. These mixed results are not surprising as we still know little about the balance between assimilation, storage and growth across temporal scales in plants [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurrence of large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
An alternative experimental approach is to reduce belowground C sink strength in tree seedlings by manipulating rooting volume, by varying the container size [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Possible advantages of this approach are that it allows a large range of treatment levels, can be easily compared to naturally planted seedlings and may mimic natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including biomass partitioning, A, water relations, nutrient uptake and respiration [@nesmith1998effect;@poorter2012pot and references therein]. Inadequate rooting volume may decrease C sink strength by progressively restricting root growth in growing plants [@thomas1991root].  Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that A is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower C exchange and assimilate export rates [@robbins1988effect]. However, evidence in support for a trade-off between C storage and growth in trees is, to date, inconclusive [@palacio2014does]. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test the extent to which growth and A are coordinated.

\
This study utilizes a novel field design to investigate the coordination between growth and A in *Eucalyptus tereticornis* Sm. seedlings, by manipulating container size and thus rooting volume. Seedlings were maintained under well watered conditions in order to isolate the effect of restricted soil volume. We used freely-rooted seedlings as a control for the container size treatments. Empirical results were combined with a simple plant growth model to simulate seedling growth with a C mass balance approach, which was then compared to observed harvested seedling mass. The model used whole-plant C gain, scaled from instantaneous rates of leaf A, to quantify seedling dry mass production over the 120 day experiment.

@@

\
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C on a random subset of each container size (n=3) after new leaves were first produced (March 13-14^th^, 2013) and prior to the final harvest (May 14-15~th~, 2013). Each AC~i~ curve was started at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l^-1^ at 25 &deg;C at saturating light (above). From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@Duursma2014] in R [@RDevelopmentCoreTeam2011].

\
Leaf dark respiration rates (R) was measured on each seedling during the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions. Reported values of R are standardized rates at 25 &deg;C using a Q10 value (1.86) developed for these seedlings in a separate experiment (Drake et al. unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

## Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and R~d~. Leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at midday 13:00-14:30 h. These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

@@

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in [@mitchell2013drought]. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, were calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass or the leaf nitrogen content. Similarly, TNC-free leaf N (N~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC and leaf N content.

## Seedling growth model
We developed a simple seedling growth model that utilized leaf A rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~i~, stem~i~ and root~i~) and a starting leaf area (LA~i~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (P~i~) is then given by

@@

\
C~day~ was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package in R with the mean photosynthetic parameters (J~max~, V~cmax~, R~d~ and g~1~) for each treatment and meteorological data from an onsite weather station. J~max~ and Vc~max~ were estimated from AC~i~ curves (explained above), R~d~ was empirically measured and the g~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] with observed g~s~ values. Methods of the coupled leaf gas exchange model are described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf A rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. C~day~ was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of C~day~ for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of C~day~ that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

\
It was further necessary to calculate a self-shading parameter ($\sigma$~s~) when scaling leaf A with total plant leaf area. This was accomplished by utilizing 61 previously digitized Eucalyptus seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package [@YplantQMC] in R to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Inputting the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total A, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf.  The ratio of total A with self-shading to horizontal leaf was then used to calculate $\sigma$~s~ for each of the 61 digitized seedlings, independently for each of the seven soil volume treatments. Next, the linear relationship between $\sigma$~s~ and total leaf area was for determined across digitized seedlings, within each treatment. For the growth model, $\sigma$~s~ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to C~day,i~. All default parameters used in model simulations are reported in Table S1.

\
We then utilized this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf A where sufficient to accurately predict overall seedling biomass production after 120 days. Each model run utilized changes in A and leaf mass fraction (LMF), with values of stem and root respiration rates, to generate total seedling mass and leaf area after 120 days.  Cumulative net leaf C gain for each treatment was equal to the sum of each value of C~day,i~ over 120 days and final seedling C was assumed to equal half of the final mass for both modeled and observed seedlings.

@@

##Data analysis
Differences in experimental parameters with soil volume were analysed by mixed-effects models in R with individual containers and experimental blocks as random effects and soil volume treatment as a categorical fixed effect with seven levels. Tukey's post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among soil volume treatments were different. A linear mixed effect model of A~max~ and leaf chemistry was performed using the 'nlme' package [@nlme] in R. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Tests of allometric relationships between log-transformed biomass components were implemented using standardized major axis regression in the 'smatr' package in R [@warton2012smatr]. Results were considered significant at P < 0.05.

@@

## Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table 3) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when foliar N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was positively correlated with final harvest biomass across all seedlings (P < 0.001).

@@

## Modelling seedling biomass
The default model M~free~, successfully optimized a LMF (21.6 %) which then allowed the model to predict mean harvest total biomass of free seedlings within 1.2 %. Using this optimized LMF, the total biomass of modeled seedlings for each soil volume treatment (M~pots~) were on average 23&pm;2.4 g C more than measured seedling biomass when compared against predicted total net leaf C gain (Figure `r massmodelfig`a). Thus, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the soil volume treatments (Figure `r massmodelfig`b). As a result, the observed reductions in leaf A with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the  container size treatments.

@@

# Discussion
This study utilized a simple but novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. We addressed questions regarding the coordination of A and growth by complementing empirical results with a C mass balance model. We found that reductions in leaf A across container sizes, when integrated across the 120 day experiment, were alone insufficient to account for observed reductions in total plant biomass production.

@@

\
Decreases in A~sat~ occurred at the same time as reductions in height and diameter of seedlings in containers. This initially suggests a strong link between growth and an apparent down regulation of A. However, there are several possible mechanisms that can explain reduced A in small containers including nutrient content, water or reduced sink strength [@poorter2012pot]. It was therefore necessary to examine each of these factors to determine if the induced belowground sink limitation actually triggered photosynthetic down regulation.

\
With high rates of g~s~, non-limiting leaf water potential and consistent leaf $\delta$^13^C across soil volume treatments there was little evidence that water stress caused the reduction in A. This finding is consistent with other container size studies without drought treatments.  For example, reduced A~max~ in cotton seedlings grown at elevated CO~2~ was attributed to sink-limited feedback inhibition from inadequate rooting volume, not decreased g~s~ [@thomas1991root]. Additionally, severe reductions in A in coffee plants were not attributed to impacts of container size on leaf water potentials or g~s~ [@ronchi2006growth]. It is likely that reductions in A of well-watered seedlings observed in our study of *E.tereticornis* seedlings was instead the result of limiting soil nutrients or space restriction on belowground sink strength.

\
Here, reductions in A were positively correlated with decreases in leaf N and leaf N was considerably reduced for seedlings in containers. As leaf N reductions were detected with TNC-free leaf mass, this suggests that physical root restriction or decreased supply likely affected seedling N uptake instead of TNC dilution in leaves in the smallest containers. Root N at the end of the experiment was on average higher in free seedlings but not consistently higher than every container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N in free seedlings, but this effect is unknown. Combined with the fact that soil N declined evenly across all treatments, there was no clear mechanism present to identify changes in root N uptake between free seedlings and seedlings in containers. In these already low quality soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N, the affects of belowground sink limitation on A of seedlings in containers was evident throughout the experiment.

>>>>

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, C allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (*A*) to allocate daily C uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in *A* explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in down regulation of *A*, these reductions were not large enough to explain observed growth responses. Thus, as *A* and growth were not tightly coordinated, excess photosynthetic C not attributed to biomass resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirms that important feedbacks exist between belowground sink strength and leaf C uptake.

@@

Understanding plant growth and its relationship to C assimilation requires knowledge of the mass balance that must be achieved between C uptake and subsequent allocation to growth, storage, and respiration. As woody plants have highly integrated systems of competing carbohydrate sinks [@kozlowski1992carbohydrate], growth should principally depend on the allocation of photosynthate among different tissues and organs. At long enough time scales photosynthesis (*A*) and respiratory losses together determine net C balance and will be coordinated to plant growth. At shorter temporal scales, however, growth can instead be mediated by tissue C storage pools. This has led to the current debate on how strongly plant growth is controlled by either source or sink activity. Consequently, plant growth cannot always be simply determined by the photosynthesis rate, making it complex to understand and challenging to model [@fourcaud2008plant]. Despite a wealth of studies, large uncertainties still remain regarding the coordination of C supply and growth of woody species.

\
In woody species, the coordination of *A* and growth has been studied with manipulations of C source activity. Examples included elevated CO~2~ experiments, for example FACE [reviewed in @ainsworth2005have], and partial defoliation experiments. Elevated CO~2~ has been shown to increase *A* [@drake1997more; @ainsworth2007response] and across four FACE experiments this resulted in a stimulation of 23 % in forest biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. In partial defoliation experiments, increases in *A* of the remaining foliage are commonly shown, yet are attributed to various mechanisms, including reduction in end product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end], increased stomatal conductance [@layne1995end], enhanced leaf nutrient status [@turnbull2007increased] and regulatory sugar signaling [@eyles2013whole]. However, increases in *A* in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], C limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant C gain [@ovaska1993a]. These manipulations of C source activity expose unresolved issues with how changes in *A* do not always infer similar responses in growth.

\
Alternatively, manipulating plant tissue C sinks is often used to investigate the correlation of *A* and growth. This is because metabolic signaling networks, relaying information on C and N status of different tissues, can regulate photosynthetic activity [@paul2001sink]. If sink inhibition of *A* occurs, a close coordination between declines in *A* and growth should be expected. Whether photosynthetic down regulation is evident in woody species has been tested through fruit removal, phloem girdling and low temperatures at high elevations. In these studies, down regulation of *A* was frequently correlated to carbohydrate accumulation resulting from reduced tissue sink strength  [@iglesias2002regulation;@hoch2002altitudinal; @urban2007girdling; @haouari2013fruit]. However, reductions in *A* were also attributed to biochemical limitations prior to carbohydrate accumulation [@nebauer2011photosynthesis], irreversible photo-oxidative damage [@duan2008photosynthetic] and stomatal limitation [@li2005photosynthesis]. These mixed results are not surprising as we still know little about the balance between assimilation, storage and growth across temporal scales in plants [@smith2007coordination]. As these manipulations likely impact source as well as sink activity simultaneously, affect water transport, are very extreme, or are specific to the occurrence of large fruiting sinks, they tell us little about source-sink coordination in typical growing conditions for woody species.

\
An alternative experimental approach is to reduce belowground C sink strength in tree seedlings by manipulating rooting volume, by varying the container size [@arp1991effects; @nesmith1998effect; @poorter2012pot]. Possible advantages of this approach are that it allows a large range of treatment levels, can be easily compared to naturally planted seedlings and may mimic natural conditions as seedlings compete for space or reach bedrock. Seedlings undergo many physiological and morphological changes in response to rooting volume, including biomass partitioning, *A*, water relations, nutrient uptake and respiration [@nesmith1998effect;@poorter2012pot and references therein]. Inadequate rooting volume may decrease C sink strength by progressively restricting root growth in growing plants [@thomas1991root].  Container size studies frequently exhibit photosynthetic down-regulation, likely as a result of sink limitation [@arp1991effects; @mcconnaughay1991physical; @gunderson1994photosynthetic; @sage1994acclimation; @maina2002intra; @ronchi2006growth]. A meta-analysis by Poorter et al. [-@poorter2012pot] concluded that *A* is the process likely to be the strongest affected by pot size and may best explain the effects on biomass seen in the large number of studies where containers are used. This conclusion arises because plants grown in small containers are shown to accumulate leaf starch while having lower C exchange and assimilate export rates [@robbins1988effect]. However, evidence in support for a trade-off between C storage and growth in trees is, to date, inconclusive [@palacio2014does]. Based on these previous studies, using container size as a sink-strength manipulation can be used to empirically test the extent to which growth and *A* are coordinated.

\
This study utilizes a novel field design to investigate the coordination between growth and *A* in *Eucalyptus tereticornis* Sm. seedlings, by manipulating container size and thus rooting volume. Seedlings were maintained under well watered conditions in order to isolate the effect of restricted soil volume. We used freely-rooted seedlings as a control for the container size treatments. Empirical results were combined with a simple plant growth model to simulate seedling growth with a C mass balance approach, which was then compared to observed harvested seedling mass. The model used whole-plant C gain, scaled from instantaneous rates of leaf *A*, to quantify seedling dry mass production over the 120 day experiment.

@@

\
Photosynthetic CO~2~ response (AC~i~) curves were measured at 25 &deg;C on a random subset of each container size (n=3) after new leaves were first produced (March 13-14^th^, 2013) and prior to the final harvest (May 14-15~th~, 2013). Each AC~i~ curve was started at the reference [CO~2~] of 400 $\mu$l l^-1^ and then consisted of 12 additional steps from [CO~2~] of 50 to 1800 $\mu$l^-1^ at 25 &deg;C at saturating light (above). From these curves the photosynthetic parameters, J~max~ and Vc~max~, were quantified using the biochemical model of [@farquhar1980biochemical] and fit with the 'plantecophys' package [@duursma2015] in R [@RDevelopmentCoreTeam2011].

\
Leaf dark respiration rates (R) was measured on each seedling during the same dates as AC~i~ curves. Freshly detached leaves were collected at least 1 hour after sundown and placed inside a conifer chamber attached to the Licor 6400. Measurements were taken at a reference [CO~2~] of 400 $\mu$l l^-1^ while leaf temperature was maintained at current ambient conditions. Reported values of R are standardized rates at 25 &deg;C using a Q~10~ value (1.86) developed for these seedlings in a separate experiment (Drake et al. unpublished). Leaf area and dry mass were recorded for each leaf during gas exchange campaigns.

## Leaf water potential
Predawn ($\Psi$~pd~) and midday ($\Psi$~l~) leaf water potentials were measured for each seedling using a PMS 1505D pressure chamber (PMS Instruments, Albany, OR, USA) on fully expanded leaves during the same time period as AC~i~ and R. Leaves were detached and immediately stored inside foil covered bags before water potential measurements were performed. $\Psi$~pd~ was measured before sunrise and $\Psi$~l~ at midday 13:00-14:30 h. These measurements were used as a measure of static water stress on the seedlings [@sellin1999does] and to ensure that the bulk soil water availability was high enough for plants to avoid water stress as they became larger and roots filled the soil volume.

@@

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in [@mitchell2013drought]. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, were calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass or the leaf N content. Similarly, TNC-free leaf N (N~f~, %) was calculated on all gas exchange leaves from leaf mass without TNC and leaf N content.

## Seedling growth model
We developed a simple seedling growth model that utilized leaf *A* rates to allocate daily C assimilate towards biomass production of stems, leaves, fine roots and coarse roots. The model begins with mean initial tissue component biomass (leaf~i~, stem~i~ and root~i~) and a starting leaf area (LA~i~) measured prior to planting. The initial biomass of roots was divided evenly between fine and coarse roots. The daily net biomass production of seedlings (P~i~) is then given by

@@

\
C~day~ was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package in R with the mean photosynthetic parameters (J~max~, V~cmax~, R and g~1~) for each treatment and meteorological data from an onsite weather station. J~max~ and Vc~max~ were estimated from AC~i~ curves (explained above), R was empirically measured and the g~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] with observed g~s~ values. Methods of the coupled leaf gas exchange model are described in Duursma et al [-@duursma2014peaked]. Combined with the meteorological parameters; PPFD, air temperature, and relative humidity, at 15 min intervals, leaf *A* rates ($\mu$mol CO~2~ m^-2^ s^-1^) were then predicted for each soil volume treatment. C~day~ was calculated by converting predicted rates to mass C gain over 15 min time steps (g m^-2^) and then summed for 24 h. This resulted in 120 unique values of C~day~ for each soil volume treatment, one value for each day of the experiment. Thus, each daily time step for model runs included a value of C~day~ that represented both treatment specific photosynthetic parameters and meteorological constraints across the duration of the experiment.

\
It was further necessary to calculate a self-shading parameter ($\sigma$~s~) when scaling leaf *A* with total plant leaf area. This was accomplished by utilizing 61 previously digitized Eucalyptus seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package [@YplantQMC] in R to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Inputting the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total *A*, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf.  The ratio of total *A* with self-shading to horizontal leaf was then used to calculate $\sigma$~s~ for each of the 61 digitized seedlings, independently for each of the seven soil volume treatments. Next, the linear relationship between $\sigma$~s~ and total leaf area was for determined across digitized seedlings, within each treatment. For the growth model, $\sigma$~s~ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to C~day,i~. All default parameters used in model simulations are reported in Table S1.

\
We then utilized this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A* where sufficient to accurately predict overall seedling biomass production after 120 days. Each model run utilized changes in *A* and leaf mass fraction (LMF), with values of stem and root respiration rates, to generate total seedling mass and leaf area after 120 days.  Cumulative net leaf C gain for each treatment was equal to the sum of each value of C~day,i~ over 120 days and final seedling C was assumed to equal half of the final mass for both modeled and observed seedlings.

@@

##Data analysis
Differences in experimental parameters with soil volume were analysed by mixed-effects models in R with individual containers and experimental blocks as random effects and soil volume treatment as a categorical fixed effect with seven levels. Tukey's post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among soil volume treatments were different. A linear mixed effect model of A~max~ and leaf chemistry was performed using the 'nlme' package [@nlme] in R. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Tests of allometric relationships between log-transformed biomass components were implemented using standardized major axis regression in the 'smatr' package in R [@warton2012smatr]. All tests of statistical significance were conducted at an alpha of 0.05.

@@

## Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table 3) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when leaf N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was positively correlated with final harvest biomass across all seedlings (P < 0.001).

@@

## Modelling seedling biomass
The default model M~free~, successfully optimized a LMF (21.6 %) which then allowed the model to predict mean harvest total biomass of free seedlings within 1.2 %. Using this optimized LMF, the total biomass of modeled seedlings for each soil volume treatment (M~pots~) were on average 23&pm;2.4 g C more than measured seedling biomass when compared against predicted total net leaf C gain (Figure `r massmodelfig`a). Thus, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the soil volume treatments (Figure `r massmodelfig`b). As a result, the observed reductions in leaf *A* with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the  container size treatments.

@@

# Discussion
This study utilized a simple but novel field design to manipulate belowground sink limitation and physically restrict *Eucalyptus tereticornis* seedling biomass production. We addressed questions regarding the coordination of *A* and growth by complementing empirical results with a C mass balance model. We found that reductions in leaf *A* across container sizes, when integrated across the 120 day experiment, were alone insufficient to account for observed reductions in total plant biomass production.

@@

\
Decreases in A~sat~ occurred at the same time as reductions in height and diameter of seedlings in containers. This initially suggests a strong link between growth and an apparent down regulation of *A*. However, there are several possible mechanisms that can explain reduced *A* in small containers including nutrient content, water or reduced sink strength [@poorter2012pot]. It was therefore necessary to examine each of these factors to determine if the induced belowground sink limitation actually triggered photosynthetic down regulation.

\
With high rates of g~s~, non-limiting leaf water potential and consistent leaf $\delta$^13^C across soil volume treatments there was little evidence that water stress caused the reduction in *A*. This finding is consistent with other container size studies without drought treatments.  For example, reduced A~max~ in cotton seedlings grown at elevated CO~2~ was attributed to sink-limited feedback inhibition from inadequate rooting volume, not decreased g~s~ [@thomas1991root]. Additionally, severe reductions in *A* in coffee plants were not attributed to impacts of container size on leaf water potentials or g~s~ [@ronchi2006growth]. It is likely that reductions in *A* of well-watered seedlings observed in our study of *E.tereticornis* seedlings was instead the result of limiting soil nutrients or space restriction on belowground sink strength.

\
Here, reductions in *A* were positively correlated with decreases in leaf N and leaf N was considerably reduced for seedlings in containers. As leaf N reductions were detected with TNC-free leaf mass, this suggests that physical root restriction or decreased supply likely affected seedling N uptake instead of TNC dilution in leaves in the smallest containers. Root N at the end of the experiment was on average higher in free seedlings but not consistently higher than every container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N in free seedlings, but this effect is unknown. Combined with the fact that soil N declined evenly across all treatments, there was no clear mechanism present to identify changes in root N uptake between free seedlings and seedlings in containers. In these already low quality soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N, the affects of belowground sink limitation on *A* of seedlings in containers was evident throughout the experiment.

\
As both rooting space and resources were finite in containers, the inability of seedlings to maintain the capacity of the belowground C sink resulted in the buildup of C assimilate in leaves. The feedback inhibition of *A* from starch accumulation has been proposed, yet it is still not known whether there is a starch threshold that triggers the down-regulation process (Nebauer et al. 2011).  Here, declines in A~max~ were correlated with higher starch content throughout the experiment. This agrees with a study on a deciduous conifer by Equiza et al. [-@equiza2006photosynthetic] where photosynthetic downregulation from reduced sink strength was correlated with starch content. As starch content in leaves of plants grown in the smallest containers was nearly double that of free seedlings in our study, this suggests the response of *A* to sink inhibition was regulated by this accumulation, as hypothesized.
